\hydrogen{} and \carbon{} Network Strategies for the European Energy System

### 2024-08-18 report/submissions/journal/paper/paper.tex
answer reviewer questions and make corresponding small text amendments

When applying a stricter \carbon{} emissions target and forcing the model to remove 460~Mt \carbon{} over one year (equivalent to 10\% of \carbon{} emissions in 1990) while sequestering maximally 660~Mt, investment in DAC and associated solar and wind power and heat supply covers a large share of additional system costs in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}). In the \hybridscenario{}, this share is at 80\% while new carbon capture at point sources only makes up 6.5\% of new system costs. This shows that the exploitation of carbon point sources is close to its feasible limit. The additional sequestration is primarily supplied by large DAC facilities near the shore. To a much smaller extent, carbon captured from bio-energetic sources in the inland, that was used to supply CU, before is now transported and sequestered (compare Fig.~\ref{fig:balance_maps_full_nn}). Some of the CU in Central Europe moves to Spain, where it uses carbon from new DAC facilities.

In the network layouts, the slight shift in CU deployment results in a partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a). This adjustment, however, leaves 80\% of the hydrogen topology intact. Simultaneously, the \carbon{} network expands routes from Central and Eastern Europe to the shores (1.2~bn€/a). The benefits of \carbon{} transport and sequestration with tighter emission targets are also evident when comparing the other scenarios. The carbon network in the \carbonscenario{} expands to transport more carbon from point sources to sequestration sites rather than to CU facilities. In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains mainly untouched (for detailed cost contributions, refer to sections~\ref{sec:cost_comparison} and~\ref{sec:operation_nn} in the Appendix).

@@

\section{Conclusion}
\label{sec:conclusion}
Our study assesses the roles of hydrogen (\hydrogen{}) and carbon dioxide (\carbon{}) networks in Europe's future energy system under net carbon neutrality and net-removal targets. We show that both networks strongly impact the optimal deployment of carbon capture, carbon utilization and sequestration. A \carbon{} network effectively reduces costs by transporting low-cost carbon from distributed point sources to sequestration sites and sites with low-cost hydrogen production to produce carbonaceous fuels. A hydrogen network provides greater cost savings as low-cost hydrogen can be transported to regions with high demand, allowing carbon utilization at point sources. A combination of both networks emphasizes the strengths of each infrastructure. In the hybrid configuration, the \hydrogen{} network supplies low-cost hydrogen to spatially fixed hydrogen demand and carbon utilization at point sources across regions. At the same time, the \carbon{} network transports low-cost carbon captured from point sources close to the coast to sequestration sites, reducing the overall reliance on high-cost Direct Air Capture. This results not only in the lowest system cost but also in a more robust energy system.

@@

The study is conducted based on the open-source, capacity expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
The model uses the open-source software tools PyPSA~\cite{brownPyPSAPythonPower2018} and Linopy~\cite{hofmannLinopyLinearOptimization2023}
for modeling and cost optimization.
The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.
All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
Many of the assumptions contained therein are based on the technology catalogs published by the Danish Energy Agency~\cite{danishenergyagencyTechnologyDataGeneration2019,thedanishenergyagencyTechnologyDataCarbon2023}.
Endogenous model results include the expansion of renewable energy sources, storage technologies, Power-To-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversion technologies.

>>>>

When applying a stricter \carbon{} emissions target and forcing the model to remove 460~Mt \carbon{} over one year (equivalent to 10\% of \carbon{} emissions in 1990) while sequestering maximally 660~Mt, investment in DAC and associated solar and wind power and heat supply covers a large share of additional system costs in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}). In the \hybridscenario{}, this share is at 80\% while new carbon capture at point sources only makes up 6.5\% of new system costs. This shows that the exploitation of carbon point sources is close to its feasible limit. The additional sequestration is primarily supplied by large DAC facilities near the shore. To a much smaller extent, carbon captured from bio-energetic sources in the inland, that under a net-zero emissions assumption supplies CU, is instead transported and sequestered (compare Fig.~\ref{fig:balance_maps_full_nn}). Some of the CU which in net-zero scenarios locates in Central Europe instead locates in Spain, where it uses carbon from new DAC facilities.

In the \carbon{} and \hydrogen{} scenarios, this slight shift in CU deployment results in a lower buildout of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a). Nonetheless, 80\% of the hydrogen topology which is built under net-zero emissions is also built under the assumption of net-negative emissions. Simultaneously, under net-negative emissions the \carbon{} network also includes routes from Central and Eastern Europe to the shores (1.2~bn€/a). The benefits of \carbon{} transport and sequestration with tighter emission targets are also evident when comparing the other scenarios. The carbon network under the net-negative \carbonscenario{} is larger than under the net-zero \carbonscenario{} and transports more carbon from point sources to sequestration sites rather than to CU facilities. In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains almost identical to the net-zero scenarios (for detailed cost contributions, refer to sections~\ref{sec:cost_comparison} and~\ref{sec:operation_nn} in the Appendix).

@@

\section{Conclusion}
\label{sec:conclusion}
Our study assesses the roles of hydrogen (\hydrogen{}) and carbon dioxide (\carbon{}) networks in Europe's future energy system under net carbon neutrality and net-removal targets. We show that both networks strongly impact the optimal deployment of carbon capture, carbon utilization and sequestration. A \carbon{} network effectively reduces costs by transporting low-cost carbon from distributed point sources to sequestration sites and sites with low-cost hydrogen production to produce carbonaceous fuels. A hydrogen network provides greater cost savings as low-cost hydrogen can be transported to regions with high demand, allowing carbon utilization at point sources. A combination of both networks emphasizes the strengths of each infrastructure. In the hybrid configuration, the \hydrogen{} network supplies low-cost hydrogen to spatially fixed hydrogen demand and carbon utilization at point sources across regions. At the same time, the \carbon{} network transports low-cost carbon captured from point sources near the coast to sequestration sites, reducing the reliance on Direct Air Capture. This results not only in the lowest system cost but also in a more robust energy system.

@@

The study is conducted based on the open-source, capacity expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
The model uses the open-source software tools PyPSA~\cite{brownPyPSAPythonPower2018} and Linopy~\cite{hofmannLinopyLinearOptimization2023}
for modeling and cost optimization.
The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.
All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}. The database applies a discount rate of 7\% for all major technologies, accounting for the weighted average costs of capital (WACC) for investments.
Many of the assumptions contained therein are based on the technology catalogs published by the Danish Energy Agency~\cite{danishenergyagencyTechnologyDataGeneration2019,thedanishenergyagencyTechnologyDataCarbon2023}.
Endogenous model results include the expansion of renewable energy sources, storage technologies, Power-To-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversion technologies.
